SmartCheck: Automatic and Efficient Counterexample Reduction and Generalization

### 2013-02-03 paper/paper.tex
Updates: paper, examples.

\newenvironment{code}{\begin{alltt}}{\end{alltt}}
%% \newenvironment{cols}{\begin{tabular}{m{3.6cm}|m{3.6cm}}{Haskell} &

@@

\todo{define variant, tag}
\todo{note constants are user-defined arguments, other simplifications}
\todo{thread args through}
\todo{termination for each algorithm}
\todo{fromMaybe over pattern matching?}
\todo{show quantification in intro}
\todo{benchmark feat}

\section{Introduction}\label{sec:intro}
The QuickCheck testing framework was a revolutionary step-forward in
type-directed random testing~\cite{qc}.  Originally designed for Haskell,
QuickCheck has been ported to other languages and is a now a widely-used testing
tool.  Because QuickCheck generates random values for testing, counterexamples
it finds may be substantially larger than a minimal counterexample.  In their
original QuickCheck paper~\cite{qc}, the authors report the following user
experience by Andy Gill:
%
\begin{quote}
Sometimes the counterexamples found are very large and it is difficult to go
back to the property and understand why it is a counterexample.
\end{quote}
%
\noindent
QuickCheck defines a type class \ttp{Arbitrary} that presents a method
\ttp{arbitrary} for generating random values of a given type.  Gill added another
method to the type class:
%
\begin{code}
smaller :: a -> [a]
\end{code}
%
\noindent
The purpose of \ttp{smaller} is to generate strictly smaller values (according to
some measure) from a given counterexample to attempt to find a smaller
counterexample.  Today, \ttp{smaller} is called \ttp{shrink}.

\begin{figure}[ht]
\begin{code}
data A = A Int16 deriving Show

data B = B [A] [A] [A] [A] deriving Show

-- QuickCheck instances
instance Arbitrary A where
  arbitrary    = liftM A arbitrary
  shrink (A i) = map A (shrink i)

instance Arbitrary B where
  arbitrary = liftM4 B arbitrary arbitrary
                       arbitrary arbitrary
  shrink (B a b c d) =
    [ B w x y z | w <- shrink a
                , x <- shrink b
                , y <- shrink c
                , z <- shrink d ]

-- Predicates for defining properties
add :: [A] -> Int16
add = sum . map (\(\backslash\)(A i) -> i)

pre :: B -> Bool
pre (B a b c d) =
  all (\(\backslash\) -> add x < 16) [a, b, c, d]

post :: B -> Bool
post (B a b c d) =
  add a + add b + add c + add d < 64

-- Property
prop :: B -> Property
prop p = pre p ==> post p
\end{code}
  \caption{QuickCheck and SmallCheck for a product-type input.}
  \label{fig:initial}
\end{figure}

While \ttp{shrink} is user-defined and can have any type-correct implementation,
it is typically defined using structural recursion for composite data types.
For example, consider the \ttp{shrink} instances given to data types \ttp{A} and
\ttp{B} in Figure~\ref{fig:initial}.\footnote{All examples and algorithms in
  this paper are presented in Haskell~\cite{haskell98}.}  For \ttp{A}, we map
the constructor \ttp{A} over a list of new \ttp{Int16}s returned by
\ttp{shrink}; for \ttp{B}, we use Haskell's list comprehension notation to
return a list of new values.  Indeed, because of the regular structure of
\ttp{shrink} implementations, the principal motivation for one of Haskell's
first generics package was for providing generic definitions for
\ttp{shrink}~\cite{syb}.

However, this straightforward approach to defining \ttp{shrink} instances may
either miss minimal counterexamples, be too inefficient to be practical, or
both.\footnote{For a recent real-world case, see the user question posed on
  StackOverflow:
  \url{http://stackoverflow.com/questions/8788542/how-do-i-get-good-small-shrinks-out-of-quickcheck}.}
For example, consider again Figure~\ref{fig:initial}.  The data type \ttp{B} is
a product type over lists of 16-bit signed integers, wrapped by a constructor.
Consider the property \ttp{prop} in the figure, stating that if the sum of each
list in a field of \ttp{B} is less than 16, then the sum of the four fields is
less than 64.  The property seems reasonable at first glance, until one realizes
that due to underflow, the property can be violated.  For example, consider the
value
%
\begin{code}
B [A (-32769)] [] [] []
\end{code}
%
\noindent
Without shrinking, QuickCheck finds large examples to the property; on average,
QuickCheck returns a value contains 70 \ttp{Int16} values!

Thus, it pays to define \ttp{shrink}.  Unfortunately, even for a simple property
and program like this one, the definition for \ttp{shrink} given in
Figure~\ref{fig:initial} produces an intractable number of potential
counterexamples to test; with the definitions and property provided (together
with QuickCheck's default \ttp{Arbitrary} instance for lists), the list of
potential counterexamples generated by \ttp{shrink} can contain more than
$10^{10}$ elements \todo{true?}.  This is an enormous state space to search for a smaller
example and can cause the shrinking phase to take orders-of-magnitude more time
than the counterexample discovery phase.

@@

Typically, a programmer tries a brute-force approach to reduce the number of
potential counterexamples; for example, she may try to truncate lists using the
(\ttp{take n}) function that returns the first $n$ elements of a list.  The
trade-off is quicker shrinking with a lower probability of finding a smaller
counterexample.  For example, redefining shrink for type \ttp{B} as follows
%

%% \begin{minipage}{\columnwidth}
%% \lstset{ caption={Truncated shrinking function for data type \ttp{B}.}
%%        , label=lst:newshrink
%%        , captionpos=b }
%% \begin{lstlisting}

\begin{figure}[ht]
\begin{code}
len = 10

shrink (B a b c d) =
  [ B w x y z | w <- tk a, x <- tk b
              , y <- tk c, z <- tk d ]
  where tk x = take len (shrink x)
\end{code}
  \caption{Truncated shrinking function for data type \ttp{B}.}
  \label{lst:newshrink}
\end{figure}

%% \end{lstlisting}
%% \end{minipage}
%
\noindent
controls the exponential blowup of the input-space by taking the first ten
elements of each list, so the cross product is limited to 10,000 elements.  The
downside is that potentially smaller counterexamples may be omitted.

@@

\begin{figure}
  \begin{code}
instance Serial Int16 where
  series d = drawnFrom [(-d')..d']
    where d' = fromIntegral d

instance Serial A where series = cons1 A
instance Serial B where series = cons4 B
  \end{code}
  \caption{SmallCheck instances for the data types in
    Figure~\ref{fig:initial}.}
  \label{fig:smallshrink}
\end{figure}

With SmallCheck, a property is checked up to a depth set by the user that
measures how large values are.  For example, for type \ttp{Int}, the depth $d$
is defined as the integers in the list enumerating from $-d$ to $d$ (i.e.,
\ttp{[(-d)..d]}).  For a product type, SmallCheck must check values produced by
taking the cross-product of the type's fields at a given depth.
Figure~\ref{fig:smallshrink} contains functions for generating Lazy SmallCheck
(a variant of SmallCheck) tests by defining instances for the \ttp{Serial}
class.  For the property \ttp{prop} defined above (suitably redefined to use
SmallCheck's types), SmallCheck must check to depth 16 to find the first
counterexample.  Unfortunately, after a couple hours of testing, Lazy SmallCheck
is still checking values at depth six, and the number of tests scales
exponentially with respect to the depth.\footnote{All tests in this paper are
  performed on a four-core Pentium~i7 running at 2.7GHz with 8GB RAM on
  Fedora~16.  This test, and others unless noted, are performed using
  interpreted Haskell under GHC 7.4.2.\todo{redo tests?}}

Finally, even if we could overcome the problems described above with QuickCheck
and SmallCheck, we are left with two others.
\begin{itemize}
  \item First, after discovering a counterexample to some property, say with
    type
%
\begin{code}
prop :: B -> Bool
\end{code}
%
\noindent
one might wish to search for different counterexamples to \ttp{prop}.  Ideally,
the testing framework would automatically add a predicate characterizing (and
perhaps generalizing) a counterexample
%
\begin{code}
cex :: B -> Bool
\end{code}
%
\noindent
as a precondition to \ttp{prop} to define a new property
\begin{code}
prop' b = not (cex b) ==> prop b
\end{code}
%
\noindent
where \ttp{==>} is QuickCheck's implication operator that discards tests that
violate the precondition.

  \item Second, and more generally, a counterexample is usually representative
    of other counterexamples, but that generalization is left to the user to
    discover for herself.  Really, what would be useful would be not just a
    minimal counterexample, but a quantified formula characterizing a set of
    counterexamples.  Such a formula has the added benefit of replacing a
    potentially large sub-term of a composite data type value with a quantifier.
\end{itemize}

@@

\paragraph{SmartCheck}
Motivated by these limitations of QuickCheck and SmallCheck, we have developed
\emph{SmartCheck}.  SmartCheck extends QuickCheck to efficiently and generically
find small counterexamples as well as generalize them as described in the
preceding paragraph.  To motivate the benefit of SmartCheck, consider the table
in Table~\ref{table:results}, giving the results for using QuickCheck and
SmartCheck on the program shown in Figure~\ref{fig:initial} (we omit SmallCheck,
since as we noted, it is not even feasible for an example like this).  In the
first three columns, the results of using QuickCheck to generate counterexamples
are shown.  The first column contains the results of using QuickCheck without
shrinking to provide a baseline.  The next two columns show the result of using
QuickCheck with definitions of \ttp{shrink} that limit the lists of fields of
\ttp{B} to 10 (Figure~\ref{lst:newshrink}) and 20 elements
(Figure~\ref{lst:newshrink}, with \ttp{len = 20}), respectively.  (With the
original definition for shrink from Figure~\ref{fig:initial}, some
counterexamples cause QuickCheck to not return during the shrinking phase after
hours of computation, so we omit these.)

@@

\item an approach to generically strengthen a property with a precondition that
  characterizes an already-discovered counterexample for the purpose of finding
  new, dissimilar counterexamples(Section~\ref{sec:generalization});

@@

\subsection{Reduction Algorithm Overview}\label{sec:reduct}
The algorithm we present for efficiently searching for new counterexamples can
be thought of as a recursive breadth-first traversal over the structure of the
current counterexample value, where the value is represented as a tree.  At each
node, during the traversal, we generate arbitrary \emph{structurally} smaller
sub-values and build a new value from that, leaving the remainder of the tree
unchanged.  Intuitively, by a \emph{structurally smaller value}, we mean one
with fewer constructors.  We continue until we reach a fixed-point.

@@

\ttp{sizedArbitrary} returns a value inside the \ttp{IO} monad since it depends
on a random seed for generation (we could pass in the generator seed, but for
simplicity, let us assume one is generated locally).  The constant
\ttp{scMaxReduce} is the maximum number of tries to discover a new
counterexample by replacing \ttp{v} in \ttp{cex} and testing it.  The result of
\ttp{pass prop cex'} for
%
\begin{code}
pass :: (a -> Property) -> a -> Bool
\end{code}
%
\noindent
holds if \ttp{cex'} satisfies the property \ttp{prop}.  The property may be a
conditional, in which case the value must pass the precondition as well as the
consequent for \ttp{pass} to return \ttp{True}.  If no failure is found, we move
to the next sub-value of \ttp{cex} and continue.  However, if a new smaller
counterexample \ttp{cex'} is found, we start a new breadth-first traversal of
\ttp{cex'}, attempting to shrink it further.  The algorithm is guaranteed to
terminate: intuitively, the measure for the function is that either the index
increases or the size of the counterexample being evaluated decreases.

@@

\begin{figure}[ht!]
  \begin{center}
    \includegraphics[scale=0.5]{Figs/cex-gen}
   \end{center}
  \caption{Counterexample generalization.}
  \label{fig:cex-gen}
\end{figure}

@@

\begin{figure}

  \caption{Shape matching algorithm.}
  \label{fig:matches}
\end{figure}

@@

\todo{shrink paper rehger pointed to ``Automatic isolation of compiler errors'' from 1994}

Closely related to SmartCheck is the recent Haskell library Feat.  Feat
enumerates algebraic data types in Haskell and indexing them
quickly~\cite{feat}.  For example, Feat returns the $10^{1000}$ value in the
enumeration of lists of Booleans in under a second.  Feat combines some of the
advantages of SmallCheck and QuickCheck:

@@

\todo{avoid complexities \url{http://stackoverflow.com/questions/14006005/idiomatic-way-to-shrink-a-record-in-quickcheck}%
%
The SYB paper:%
  \url{http://research.microsoft.com/en-us/um/people/simonpj/Papers/hmap/index.htm}%
  Has shrink example%
%
* QuickSpec:%
  \url{http://www.cse.chalmers.se/~nicsma/quickspec.pdf}%
%
\url{http://blog.regehr.org/archives/697} --- check out C-reduce (Regehr) and Delta%
  debugging (\url{http://www.st.cs.uni-saarland.de/dd/}).  C-reduce is in PLDI.%
%
* Also check out \url{http://www.whyprogramsfail.com/}%
%
* GenCheck \url{https://github.com/JacquesCarette/GenCheck}%
%
* Dart-check: \url{http://code.google.com/p/dart-immutable/} (and G+ post by Paul%
  Brauner)%
}

>>>>

\newenvironment{code}{\begin{alltt}\small}{\end{alltt}}
%% \newenvironment{cols}{\begin{tabular}{m{3.6cm}|m{3.6cm}}{Haskell} &

@@

\todo{complexity of each algorithm}
\todo{termination for each algorithm}
\todo{show quantification in intro}
\todo{benchmark feat -- send to author}

\section{Introduction}\label{sec:intro}
The QuickCheck testing framework was a revolutionary step-forward in
type-directed random testing~\cite{qc}.  Originally designed for Haskell,
QuickCheck has been ported to other languages and is a now a widely-used testing
tool.  Because QuickCheck generates random values for testing, counterexamples
it finds may be substantially larger than a minimal counterexample.  In their
original QuickCheck paper~\cite{qc}, the authors report the following user
experience by Andy Gill:
%
\begin{quote}
Sometimes the counterexamples found are very large and it is difficult to go
back to the property and understand why it is a counterexample.
\end{quote}
%
\noindent
QuickCheck defines a type class \ttp{Arbitrary} that presents a method
\ttp{arbitrary} for generating random values of a given type.  Gill added another
method to the type class:
%
\begin{code}
smaller :: a -> [a]
\end{code}
%
\noindent
The purpose of \ttp{smaller} is to generate strictly smaller values (according
to some measure) from a given counterexample.  These new values are then tested
to attempt to find a smaller counterexample.  Today, \ttp{smaller} is called
\ttp{shrink}.

\begin{figure}[ht]
\begin{code}
data N a = N a
type I   = [N Int16]
data T   = T (N Word8) I I I I

toList :: T -> [[Int16]]
toList (T w i0 i1 i2 i3) =
  [go w] : (map . map) go [i0, i1, i2, i3]
  where go (N i) = fromIntegral i

pre :: T -> Bool
pre t = all ((>) 256 . sum) (toList t)

post :: T -> Bool
post t = (sum . concat) (toList t) < 5 * 256

prop :: T -> Property
prop t = pre t ==> post t
\end{code}
  \caption{QuickCheck and SmallCheck for a product-type input.}
  \label{fig:initial}
\end{figure}

While \ttp{shrink} is user-defined and can have any type-correct implementation,
it is typically defined using structural recursion for composite data types.
Indeed, because of the regular structure of \ttp{shrink} implementations, the
principal motivation for one of Haskell's first generics package was for
providing generic definitions for \ttp{shrink}~\cite{syb}.  However, simple
structural recursion may either miss smaller counterexamples, be too inefficient
to be practical, or both.\footnote{For a recent real-world case, see the user
  question posed on StackOverflow:
  \url{http://stackoverflow.com/questions/8788542/how-do-i-get-good-small-shrinks-out-of-quickcheck}.}

Consider the example in Figure~\ref{fig:initial}.\footnote{All examples and
  algorithms in this paper are presented in Haskell.  We strive to use idiomatic
  Haskell, and only rely on the language extensions of existential type
  quantification and scoped type variables.~\cite{haskell98}.}  Data type
\ttp{T} is a product type containing four lists of \ttp{Int16}s and a single
\ttp{Word8}.  For our purposes, it is not important what \ttp{T} models, but
for concreteness, suppose it models the values in four linked lists in an
operating system, together with an 8-bit status register.

Now suppose we are modeling some program that serializes values of type \ttp{T}.
The input to the program satisfies the invariant \ttp{pre}, that the sum of
values in each list of \ttp{Int16}s is less than or equal to 256 (and by
definition, the \ttp{Word8} value is less than or equal to 256).  Assuming this,
we want to show \ttp{post} holds, that the sum of all the values from \ttp{T} is
less than $5 * 256$.  At first glance, the property seems reasonable, but we
have forgotten about underflows.  For example, the value
%
\begin{code}
T (N 10) [N (-20000)] [N (-20000)] [] []
\end{code}
%
fails \ttp{prop} (\ttp{==>} is implication from the QuickCheck library).
Unfortunately, a typical counterexample returned by QuickCheck is not quite so
simple.  With standard settings, it returns a value containing just under 70
\ttp{N} constructors, and sometimes values contain over 100 constructors.  Thus,
it pays to define \ttp{shrink}!

A typical definition of \ttp{shrink} for a data type like \ttp{T} is to take the
cross-product of shrunk values over the arguments to the constructor \ttp{T}.
This can be expressed using Haskell's list-comprehension notation:
%
\begin{code}
shrink (T w i0 i1 i2 i3) =
  [ T a b c d e | a <- shrink w,  b <- shrink i0
                , c <- shrink i1, d <- shrink i2
                , e <- shrink i3 ]
\end{code}
%
\noindent
The result of (\ttp{shrink t}) with this definition is null if any list
contained in \ttp{t} is null.  Still, for typical counterexamples returned by
QuickCheck that do not contain empty lists, together with QuickCheck's default
\ttp{Arbitrary} instances, the length of potential counterexamples returned by
\ttp{shrink} can easily exceed $10^{10}$, which is an intractable number of
tests for QuickCheck to analyze.  The time complexity is compounded by the
shrink stage being recursive, so when a new smaller counterexample is found, it
is also shrunk.  In practice, QuickCheck appears to loop indefinitely.

@@

Typically, a programmer tries a brute-force approach to reduce the number of
potential counterexamples; for example, she may try to truncate lists using the
(\ttp{take n}) function that returns the first $n$ elements of a list.  The
trade-off is quicker shrinking with a lower probability of finding a smaller
counterexample.  For example, we might redefine shrink as follows:
%
%\begin{figure}[ht]
\begin{code}
len = 10

shrink (T w i0 i1 i2 i3) =
  [ T a b c d e | a <- tk w
                , b <- tk i0, c <- tk i1
                , d <- tk i2, e <- tk i3 ]
  where tk x = take len (shrink x)
\end{code}
%%   \caption{Truncated shrinking function for data type \ttp{B}.}
%%   \label{lst:newshrink}
%% \end{figure}

%% \end{lstlisting}
%% \end{minipage}
%
\noindent
let us call this version ``truncated shrink''.  Truncating the controls the
blowup of the input-space; the number of possible new values is $10^5$ in this
case.  While truncation controls the blowup, the downside is that potentially
smaller counterexamples may be omitted.

@@

%% \begin{figure}
%%   \begin{code}
%% instance Serial Int16 where
%%   series d = drawnFrom [(-d')..d']
%%     where d' = fromIntegral d

%% instance Serial A where series = cons1 A
%% instance Serial B where series = cons4 B
%%   \end{code}
%%   \caption{SmallCheck instances for the data types in
%%     Figure~\ref{fig:initial}.}
%%   \label{fig:smallshrink}
%% \end{figure}

With SmallCheck, a property is checked up to a depth set by the user that
measures how large values are (typically, the depth is a number of constructors
in the value).  For a product type, SmallCheck must check values produced by
taking the cross-product of the type's fields at a given depth.  SmallCheck is
unable to enumerate all possible values of type \ttp{T} to discover a
counterexample, due to combinatorial explosion.  (A related library named Feat
combines some aspects of SmallCheck and QuickCheck~\cite{feat}; we discuss it
more in the related work (Section~\ref{sec:related}).

%% is unable to verify the property from Figure~\ref{fig:initial}, for much the
%% same reason that \ttp{shrink} is inefficient

%% For the property \ttp{prop} defined in Figure~\ref{fig:initial} (suitably
%% redefined to use SmallCheck's types).

%% to depth 16 to find the first counterexample.  Unfortunately, after a couple
%% hours of testing, Lazy SmallCheck is still checking values at depth six, and
%% the number of tests scales with respect to the depth.\footnote{All tests in
%% this paper are performed on a four-core Pentium~i7 running at 2.7GHz with 8GB
%% RAM on Fedora~16.  This test, and others unless noted, are performed using
%% interpreted Haskell under GHC 7.4.2.\todo{redo tests?}}

%% For example, for type \ttp{Int}, the depth $d$ is defined as the integers in
%% the list enumerating from $-d$ to $d$ (i.e., \ttp{[(-d)..d]}).
%% Figure~\ref{fig:smallshrink} contains functions for generating Lazy
%% SmallCheck (a variant of SmallCheck) tests by defining instances for the
%% \ttp{Serial} class.

Finally, even if we could overcome the problems described above with QuickCheck
and SmallCheck, we are left with two others.
\begin{itemize}
  \item A counterexample is usually representative of other counterexamples, but
    that generalization is left to the user to discover for herself.  For
    example, consider the counterexample presented earlier:
%
\begin{code}
T (N 10) [N (-20000)] [N (-20000)] [] []
\end{code}
%
\noindent
For any \ttp{Word8}, the counterexample still holds.  Thus, we can be more
informative and return a quantified formula rather than a single counterexample:
%
\begin{code}
forall x . T x [N (-20000)] [N (-20000)] [] []
\end{code}
%
\noindent
Such a formula has the added benefit of replacing a potentially large sub-term
of a composite data type value with a quantifier, reducing the sheer size of the term.

  \item Second, during testing, it is sometimes helpful to cover the input space
    and by discovering different counterexamples to \ttp{prop}.  In the example
    in Figure~\ref{fig:initial}, one counterexample is fairly representative of
    others, but this is not true in general (e.g., consider properties about
    abstract syntax trees).  Ideally, the testing framework would automatically
    add a predicate characterizing a counterexample as the precondition to
    \ttp{prop} that strengthens the property.  That is, for a specific
    counterexample \ttp{ex}, we have a predicate \ttp{predEx} that holds if its
    argument is another counterexample that is similar to (i.e., some
    generalization of) \ttp{ex}.  Then we test against a new property
%
\begin{code}
prop' a = not (cex a) ==> prop a
\end{code}
%
%% \noindent
%% where \ttp{==>} is QuickCheck's implication operator that discards tests that
%% violate the precondition.
\end{itemize}

@@

\paragraph{SmartCheck}
Motivated by these limitations of QuickCheck and SmallCheck, we have developed
\emph{SmartCheck}.  SmartCheck extends QuickCheck to efficiently and generically
find small counterexamples as well as generalize them as described in the
preceding paragraph.  To motivate the benefit of SmartCheck, consider
Table~\ref{table:results}, giving the results for using QuickCheck and
SmartCheck on the program shown in Figure~\ref{fig:initial} (we omit SmallCheck,
since as we noted, it is not even feasible for an example like this).  In the
first three columns, the results of using QuickCheck to generate counterexamples
are shown.  The first column contains the results of using QuickCheck without
shrinking to provide a baseline.  The next two columns show the result of using
QuickCheck with definitions of truncated \ttp{shrink} that limit the number of
smaller values to 10 and 20 elements, respectively.  (With the original
definition for shrink from Figure~\ref{fig:initial}, some counterexamples cause
QuickCheck to not return during the shrinking phase after hours of computation,
so we omit these.)

@@

\item an approach to generically strengthen a property with a precondition that
  characterizes an already-discovered counterexample for the purpose of finding
  new, dissimilar counterexamples (Section~\ref{sec:generalization});

@@

\subsection{Reduction Algorithm Overview}\label{sec:reduct}
The algorithm we present for efficiently searching for new counterexamples is an
instance of greedy breadth-first search over a tree structures, representing a
value.  At each node, during the traversal, we generate arbitrary
\emph{structurally} smaller sub-values and build a new value from that, leaving
the remainder of the tree unchanged.  Intuitively, by a \emph{structurally
  smaller value}, we mean one with fewer constructors.  We continue until we
reach a fixed-point.

@@

\ttp{sizedArbitrary} returns a value inside the \ttp{IO} monad since it depends
on a random seed for generation (we could pass in the generator seed, but for
simplicity, let us assume one is generated locally).  The constant
\ttp{scMaxReduce} is the maximum number of tries to discover a new
counterexample by replacing \ttp{v} in \ttp{cex} and testing it.  The result of
\ttp{pass prop cex'} for
%
\begin{code}
pass :: (a -> Property) -> a -> Bool
\end{code}
%
\noindent
holds if \ttp{cex'} satisfies the property \ttp{prop}.  The property may be a
conditional, in which case the value must pass the precondition as well as the
consequent for \ttp{pass} to return \ttp{True}.  If no failure is found, we move
to the next sub-value of \ttp{cex} and continue.  However, if a new smaller
counterexample \ttp{cex'} is found, we start a new breadth-first traversal of
\ttp{cex'}, attempting to shrink it further.  The algorithm is guaranteed to
terminate: intuitively, the measure for the function is that either the index
increases or the size of the counterexample being evaluated decreases.  The
complexity of the algorithm is $\Omega(n^2)$, where $n$ is the number of
constructors in the counterexample.  The analysis assumes generating new
sub-values and testing them is done in constant time.

@@

\begin{figure}[ht!]
  \begin{center}
    \includegraphics[scale=0.49]{Figs/cex-gen}
   \end{center}
  \caption{Counterexample generalization.}
  \label{fig:cex-gen}
\end{figure}

@@

\begin{figure}
  \begin{code}
matchesShapes :: SubTypes a => a -> [(a,([Idx],[Idx]))] -> Bool
matchesShapes d = any (matchesShape d)

matchesShape :: SubTypes a => a -> (a, ([Idx],[Idx])) -> Bool
matchesShape a (b, (idxVals, idxConstrs))
  | constr a /= constr b = False
  | Just a' <- aRepl         = let x = subForest (subVals a') in
                               let y = subForest (subVals b) in
                               all foldEqConstrs (zip x y)
  | otherwise                = False

  where
  foldEqConstrs :: (Tree SubVal, Tree SubVal) -> Bool
  foldEqConstrs (Node (SubVal l0) sts0, Node (SubVal l1) sts1)
    | constr l0 == constr l1 = next
    | otherwise              = False
    where next = all foldEqConstrs (zip sts0 sts1)

  updateA idx d = fmap (replace d idx) (index b idx)
  aRepl = foldl go (Just a) (idxVals ++ idxConstrs)
    where go ma idx = maybe Nothing (updateA idx) ma
  \end{code}
  \caption{Shape matching algorithm.}
  \label{fig:matches}
\end{figure}

@@

\todo{shrink paper rehger pointed to ``Automatic isolation of compiler errors''
  from 1994}
\todo{\url{http://www.cs.utah.edu/~regehr/papers/pldi12-preprint.pdf}}
\url{http://blog.regehr.org/archives/697} --- check out C-reduce (Regehr) and Delta%
  debugging (\url{http://www.st.cs.uni-saarland.de/dd/}).  C-reduce is in PLDI.%

Related to SmartCheck is the recent Haskell library Feat~\cite{feat}.  Feat
provides automated enumerations of algebraic data types in Haskell, allowing for
fast access to very large indexes.  For example, from the enumeration of
(\ttp{[Bool]})
%
\begin{code}
[[],[False],[True],[False,False],[False,True] ...
\end{code}
%
\noindent
Accessing the $10^{1000}$th element takes under $0.1$ seconds in interpreted
Haskell.  Feat combines some advantages of SmallCheck and QuickCheck, since the
user can choose to exhaustively test an enumeration up to some depth, like with
SmallCheck, or she can create a uniform distribution of test cases up to some
depth.

combines some of the advantages of SmallCheck and QuickCheck:

@@

\todo{avoid complexities \url{http://stackoverflow.com/questions/14006005/idiomatic-way-to-shrink-a-record-in-quickcheck}%
%
The SYB paper:%
  \url{http://research.microsoft.com/en-us/um/people/simonpj/Papers/hmap/index.htm}%
  Has shrink example%
%
* QuickSpec:%
  \url{http://www.cse.chalmers.se/~nicsma/quickspec.pdf}%
%
%
* Also check out \url{http://www.whyprogramsfail.com/}%
%
* GenCheck \url{https://github.com/JacquesCarette/GenCheck}%
%
* Dart-check: \url{http://code.google.com/p/dart-immutable/} (and G+ post by Paul%
  Brauner)%
}
%
\url{http://www.haskell.org/haskellwiki/Research_papers/Testing_and_correctness}
